---
title: "Icons, logos and annotations with zusvg"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Icons, logos and annotations with zusvg}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 5, fig.height = 2.5)
```

zusvg draws SVG icons and logos to pixels with two small C libraries
compiled into the package, so it needs no system library. It draws shapes,
paths, strokes, gradients, `<use>` and `<symbol>`, embedded PNG and JPEG
images, and the colours you give it; it does not draw text, filters, masks,
patterns, markers, CSS style sheets or clip paths. Those are what icons
rarely use and plots and diagrams often do: for those, use `rsvg` or
`magick`.

```{r setup}
library(zusvg)
```

## An icon

`svg_render()` takes the document's text, a raw vector, a connection, or a
document loaded once with `svg_load()`, and returns a `nativeRaster` that
base graphics and grid draw directly.

```{r icon}
house <- '<svg xmlns="http://www.w3.org/2000/svg" viewBox="0 0 24 24"
  fill="none" stroke="currentColor" stroke-width="2" stroke-linecap="round"
  stroke-linejoin="round"><path d="M3 10.5 12 3l9 7.5V20a1 1 0 0 1-1 1h-5v-6H9v6H4a1 1 0 0 1-1-1z"/></svg>'
r <- svg_render(house, width = 96, color = "steelblue")
dim(r)
op <- par(mar = c(0, 0, 0, 0))
plot.new()
rasterImage(r, 0.3, 0, 0.7, 1)
par(op)
```

`color` is what `currentColor` means, which is how most icon sets are
recoloured.

## An icon sheet

A sheet of icons with an `id` each is cut into icons with `svg_elements()`
and `svg_render(id = )`, which crops to the element's extents.

```{r sheet}
sheet <- '<svg xmlns="http://www.w3.org/2000/svg" width="72" height="24">
  <g id="dot"><circle cx="12" cy="12" r="9" fill="var(--a, tomato)"/></g>
  <g id="square"><rect x="27" y="3" width="18" height="18" rx="3" fill="var(--b, gold)"/></g>
  <g id="triangle"><path d="M60 3l9 18H51z" fill="var(--a, tomato)"/></g>
</svg>'
svg_elements(sheet)
icons <- lapply(svg_elements(sheet)$id, function(id) svg_render(sheet, id = id, width = 48))
op <- par(mar = c(0, 0, 0, 0))
plot.new()
for (i in seq_along(icons)) rasterImage(icons[[i]], (i - 1) / 3 + 0.05, 0.1, i / 3 - 0.05, 0.9)
par(op)
```

## Recolouring with a palette

`var(--name)` in a fill or stroke is answered from `palette`, a named
character vector, so one sheet serves several themes.

```{r palette}
dark <- svg_render(sheet, width = 216, palette = c(a = "#7fdbca", b = "#c792ea"),
                   background = "#1e1e2e")
op <- par(mar = c(0, 0, 0, 0))
plot.new()
rasterImage(dark, 0, 0.2, 1, 0.8)
par(op)
```

## A ggplot2 annotation

The `nativeRaster` becomes a grob with `grid::rasterGrob()`, and
`ggplot2::annotation_custom()` places it in data coordinates.

```{r ggplot, eval = requireNamespace("ggplot2", quietly = TRUE)}
library(ggplot2)
logo <- grid::rasterGrob(svg_render(house, width = 64, color = "grey30"))
ggplot(mtcars, aes(wt, mpg)) +
  geom_point() +
  annotation_custom(logo, xmin = 4.6, xmax = 5.4, ymin = 29, ymax = 34)
```

## Files

`svg_png()` and `svg_jpeg()` encode a render, returning bytes or writing a
file. The same document and arguments give the same bytes on every
platform. A Shiny app can serve them from `renderImage()` or embed them as a
data URI:

```{r png}
png <- svg_png(house, width = 32)
png[1:4]
path <- tempfile(fileext = ".png")
svg_png(house, path, width = 32)
file.size(path)
```

```r
# In a Shiny server function:
output$icon <- renderImage({
  path <- tempfile(fileext = ".png")
  svg_png(house, path, width = 64, color = input$colour)
  list(src = path, contentType = "image/png", width = 64, height = 64)
}, deleteFile = TRUE)
```

## Untrusted input

An SVG from a web page or a message is untrusted. `svg_load()` refuses
input over `max_size`, documents with more than `max_elements` elements or
nested deeper than `max_depth`, and, with `images = FALSE`, any embedded
image, before the renderer sees a byte; `svg_render()` refuses surfaces over
`max_pixels`, and a render stops after a million element visits. Each
refusal is a classed condition (see `?zusvg-conditions`); `svg_validate()`
returns it instead of raising it, for gating a batch.

```{r untrusted}
svg_validate('<svg xmlns="http://www.w3.org/2000/svg"><image href="x.png"/></svg>',
             images = FALSE)
```

## What does not render

Text and clip paths warn once per call, with a count, so a document that
needs them is noticed:

```{r warnings}
label <- '<svg xmlns="http://www.w3.org/2000/svg" width="40" height="20">
  <rect width="40" height="20" fill="#eee"/><text x="4" y="14">Hi</text></svg>'
r <- tryCatch(svg_render(label), zusvg_text_skipped = function(w) conditionMessage(w))
r
```
